---
title: "Book Recommender System"
author: "Tino Muzambi"
format: 
  html:
    embed-resources: true
    page-layout: full
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
# Libraries
install.if.missing <- function(package.name) {
  if (!require(package.name, character.only = TRUE)) {
    install.packages(package.name)
    library(package.name, character.only = TRUE)
  }
}

install.if.missing("tidyverse")
```

# The Data
The data we will be working with is a partially preprocessed version of the ”Book-Crossing” dataset. The data can be downloaded [here](https://www.kaggle.com/datasets/arashnic/book-recommendation-dataset/) from Kaggle. It was collected by Cai-Nicolas Ziegler in a 4-week crawl (August / September 2004) from the Book-Crossing community with kind permission from Ron Hornbaker, CTO of Humankind Systems. It contains 278 858 users (anonymised but with demographic information) providing 1 149 780 ratings (explicit / implicit) about 271 379 books.

```{r}
# Load the data
ratings <- read.csv("data/Ratings.csv")
books <- read.csv("data/Books.csv")
users <- read.csv("data/Users.csv")

head(ratings)
head(books)
head(users)
```


